---
name: Josh
surname: Whelan
position: "Choice Modeling and Data Analytics"
address: "Toronto, ON"
phone: "416-938-4834"
email: "jrwhelan@me.com"
linkedin: joshwhelan

docname: "Resume"
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::awesomecv
headcolor: 414141

profilepic: "DSC00454.JPG"
#www: joshwhelan.ca
#github: tallguyjenks
keywords: RMarkdown, Rmd, Resume

fontfamily: mathpazo
fontfamilyoptions: sc, osf
fontsize: 12pt
linkcolor: blue
urlcolor: blue

aboutme: "I like to break down problems to build pragmatic solutions. Coming from a background of economics and mathematics, I enjoy the challenge of creating well designed experiments and models to test hypotheses and produce informative analytics tools to forecast possibilities."
---


```{r setup, include=FALSE}
library(tidyverse)
library(vitae)
knitr::opts_chunk$set(echo = FALSE)
```

<!-- # Professional Goal -->

<!-- GOAL TEXT WOULD GO HERE -->

# Skills

## Data Science

* Discrete choice models, data wrangling, hierarchical Bayesian modeling, regression analysis, process automation, dynamic reports, survey analysis, classification


## Languages & Applications

* Expertise with R: tidyverse and tidy data principles, Shiny applications, Rmarkdown reports
  * Extensive experience in market research applications: MaxDiff, Conjoint, Virtual Shopping exercises
 <!--   * This resume was built with Rmarkdown and the `vitae` package -->
* Expertise with MS Excel: VBA, share of preference and volumetric simulators
* Experience with Python, SQL, AWS

<!--
## Leadership

* Managing and mentoring junior team members
* Adept communicator with experience presenting complex concepts to both technical and non-technical audiences
-->

# Experience
```{r}
tribble(
  ~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Modeling and analyzing consumer preferences through hierarchical Bayes estimation with both R and Sawtooth Software",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Creating informative simulators in Excel and R from random utility maximization models to forecast preferences towards new consumer products or concepts",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Creating and testing experimental designs for quantitative survey methods in a variety of frameworks including MaxDiff, choice-based conjoint (CBC), and full virtual market discrete choice",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Collaborating with government agencies and nonprofit organizations to lead quantitative research and conduct evaluation of programs in social return of investment contexts",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Managing and mentoring junior team members and planning around multiple agile projects with competing deadlines",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Building tools to reduce outsourcing of manual tasks or to automate processes and enable research teams to work more independently, save time, and work more efficiently",

  "Choice Modeling and Data Analytics Manager", "Dig Insights", "Toronto, ON", "May 2018 - Present", "Developing and testing new methodologies for choice modeling projects",

  "Research Assistant", "McMaster University", "Hamilton, ON", "Sep 2015 - Sep 2018", "Collaborated with professors in research projects, writing on findings and using Stata and MS Excel to clean and analyze data for empirical studies",

  "Research Assistant", "McMaster University", "Hamilton, ON", "Sep 2015 - Sep 2018", "Cleaned and restructured datasets secured in the Public Economics Data Analysis Laboratory to generate reports on confidential data for clients",

#  "Research Assistant", "McMaster University", "Hamilton, ON", "Sep 2015 - Sep 2018", "Studied advanced statistical methods and econometrics in the Department of Economics",

  "Online Product Analyst", "Thomson Reuters", "Toronto, ON", "Aug 2014 - Sep 2015", "Communicated and coordinated between management and development teams, ensuring teams were aligned on goals and capabilities",

  "Online Product Analyst", "Thomson Reuters", "Toronto, ON", "Aug 2014 - Sep 2015", "Performed QA testing for active and upcoming online legal research products and prioritizing bugs for the development team",

  "Junior Business Analyst", "Catlin Canada", "Toronto, ON", "Dec 2013 - Aug 2014", "Developed relational database solutions with Filemaker Pro and SQL to identify and analyze trends and opportunities",

  "Junior Business Analyst", "Catlin Canada", "Toronto, ON", "Dec 2013 - Aug 2014", "Designed, tested, and debugged user interfaces for in-house databases and reports",

 # "Junior Business Analyst", "Catlin Canada", "Toronto, ON", "Dec 2013 - Aug 2014", "Assisted in project planning and organization within a team of business analysts"

  #"Teaching Assistant", "York University", "Toronto, ON", "Sep 2012 - Sep 2013", "Taught seminars and held office hours to teach and tutor economic principles to students both in groups and individually",

  #"Management Analyst", "Government of Newfoundland and Labrador", "St. John's, NL", "2012", "Created a database of geospatial data on moose vehicle collisions (MVC) in the province to aid policy decisions using MS Excel and Google Earth",

  #"Management Analyst", "Government of Newfoundland and Labrador", "St. John's, NL", "2012", "Strengthened the existing provincial MVC data by adding uncertainty measures through analysis of raw data sources"
 ) %>% 
  detailed_entries(role, dates, company, loc, details)
  #detailed_entries(what, when, with, where, why)
  #brief_entries(what, when, with)

```


# Education
```{r}
tribble(
  ~ degree, ~ uni, ~ loc, ~ dates, ~ details,
  "MA, Economics", "York University", "Toronto, ON", "2012 - 2013","Research Assistant",

  "BSc Joint Honours, Economics & Applied Mathematics", "Memorial University of Newfoundland", "St. John's, NL", "2007 - 2012","Ernest Joseph Scholarship"
) %>% 
  detailed_entries(degree, dates, uni, loc)
```

