---
title: "Basketball Prediction"
---

## Introduction

Basketball games in the NBA are nice for a logistic regression approach because they only have two possible outcomes: you win or you lose. We dont have to deal with draws, even in the regular season.

We can use a individual team effects in a logistic regression approach to estimate the likelihood that a specific team gets a win against another given team. Clearly stated, we might ask based on the last 3 seasons of games, what's the probability that the Toronto Raptors beat in the Golden State Warriors for a game played in the Bay Area?


----

## Data

The [`nbastatR`](http://asbcllc.com/nbastatR/index.html) package for `R` is a conventient API wrapper to pull rich player and game statistics from the NBA Stats API and Basketball-Reference (as well as other sources for non-NBA basketball leagues). You can install the package through github with devtools.

We'll load in the team and player reference tables, as well as game data for the past 3 seasons. We'll then transform the data structure somewhat in preparation for analysis. To starts, let's create dummy variables of the winners and losers for each game.


```{r dataload, echo=FALSE, message=F}
library(nbastatR)
library(tidyverse)
library(lubridate)

today <- today(tzone = "EST")
historyCut <- year(today)-2

playerRef <- nba_players()
teamRef <- nba_teams()

teamCodes <- teamRef %>%
	dplyr::select(slugTeam, idTeam)

playerLogs <- game_logs(
  seasons = 2019,
  league = "NBA",
  result_types = "player",
  season_types = "Regular Season",
  nest_data = F,
  cols = c(dataTables),
  assign_to_environment = TRUE,
  return_message = TRUE
)

teamLogs <- game_logs(
  seasons = historyCut:year(today),
  league = "NBA",
  result_types = "team",
  season_types = "Regular Season",
  nest_data = F,
  cols = c(dataTables),
  assign_to_environment = TRUE,
  return_message = TRUE
) %>%
	dplyr::filter(locationGame=="H")


winners <- teamLogs %>%
    mutate(n=1) %>%
    pivot_wider(
      id_cols = idGame,
      names_from = slugTeamWinner,
      names_prefix = "win_",
      values_from = n,
      values_fill = list(n = 0)
    )


losers <- teamLogs %>%
    mutate(n=1) %>%
    pivot_wider(
      id_cols = idGame,
      names_from = slugTeamLoser,
      names_prefix = "lose_",
      values_from = n,
      values_fill = list(n = 0)
    )

winloss <- winners %>%
	left_join(losers)

```

----


## The Model

Hierarchical Bayesian Logistic Regression

```{r modelling}
library(RSGHB)
library(glue)

choicedata <- winloss %>%
	rename(ID = idGame) %>%
	arrange(ID)

teamnames <- teamLogs %>%
	dplyr::select(slugTeam) %>%
	dplyr::distinct()

teamList <- teamnames$slugTeam

# ---------------------------------------------------------------------------
# THE LIKELIHOOD FUCNTION

fixed <- FALSE

if (fixed==TRUE) {
# FIXED PARAMS
n <- ""
for (i in 1:length(teamList)){
  n <- glue('{n} B_{teamList[i]} <- fc[cc]; cc <- cc + 1;')
}
} else {
# RANDOM PARAMS
n <- ""
for (i in 1:length(teamList)){
	n <- glue('{n} B_{teamList[i]} <- b[, cc]; cc <- cc + 1;')
}
}

g1 <- glue('v1 <- ')
for (i in 1:(length(teamList)-1)){
	g1 <- glue('{g1} B_{teamList[i]}*choicedata$win_{teamList[i]} + ')
}
g1 <- glue('{g1} B_{teamList[length(teamList)]}*choicedata$win_{teamList[length(teamList)]} ')

g2 <- glue('v2 <- ')
for (i in 1:(length(teamList)-1)){
	g2 <- glue('{g2} B_{teamList[i]}*choicedata$lose_{teamList[i]} + ')
}
g2 <- glue('{g2} B_{teamList[length(teamList)]}*choicedata$lose_{teamList[length(teamList)]} ')

j <- glue('p <- (exp(v1)*choicedata$choice1 + exp(v2)*choicedata$choice2) / (exp(v1) + exp(v2))')
z <- glue('likelihood <- function(fc, b) {{ cc <- 1; {n} {g1}; {g2}; {j}; return(p) }}')
eval(parse(text=z))
# ---------------------------------------------------------------------------


# ITERATION SETTINGS
gNCREP    <- 50000  # Number of iterations to use prior to convergence
gNEREP    <- 2000  # Number of iterations to keep for averaging after convergence has been reached
gNSKIP    <- 10   # Number of iterations to do in between retaining draws for averaging
gINFOSKIP <- 1000    # How frequently to print info about the iteration process

iters <- gNCREP + (gNEREP*gNSKIP)



if (fixed==TRUE) {
# FIXED PARAMS
modelname <- "simpleCompare-aggregate" 
gVarNamesFixed <- glue("B_{teamList}")
FC <- rep(0, length(teamList))         # for the fixed coefficients

control <- list(
  modelname = modelname,
  gVarNamesFixed = gVarNamesFixed,
  FC = FC,
  gNCREP = gNCREP,
  gNEREP = gNEREP,
  gNSKIP = gNSKIP,
  gINFOSKIP = gINFOSKIP,
  targetAcceptanceNormal = 0.234,
  gSeed = 1984
)
} else {
# RANDOM PARAMS
modelname <- "simpleCompare" 
gVarNamesNormal <- glue("B_{teamList}")
gDIST <- c(rep(1, length(teamList)))
svN <- rep(0, length(teamList))     # for the random coefficients

control <- list(
  modelname = modelname,
  gVarNamesNormal = gVarNamesNormal,
  svN = svN,
  gDIST = gDIST,
  gNCREP = gNCREP,
  gNEREP = gNEREP,
  gNSKIP = gNSKIP,
  gINFOSKIP = gINFOSKIP,
  targetAcceptanceNormal = 0.234,
  gSeed = 1984
)
}


#HIT RATE TESTING:
#	HIT.TEST <- dplyr::filter(choicedata, Task==5)
#	write.csv(HIT.TEST,"RSGHB hit_test - noOil - negNormP.csv", row.names = FALSE)


#filtering obs:
# choicedata <- dplyr::filter(choicedata, Task!=5)

# Estimate the model
# model <- doHB(likelihood, choicedata, control)

# Save model object
######## save(model, file = paste0(model$modelname, ".RData"))

# write the model utilities
# write.csv(model$C,"RSGHB utilities - Best - negNormP -censored.csv", row.names = FALSE)



```


