---
title: "Basketball Prediction"
---

## Introduction

Basketball games in the NBA are nice for a logistic regression approach because they only have two possible outcomes: you win or you lose. We dont have to deal with draws, even in the regular season.

We can use a individual team effects in a logistic regression approach to estimate the likelihood that a specific team gets a win against another given team. Clearly stated, we might ask based on the last 3 seasons of games, what's the probability that the Toronto Raptors beat in the Golden State Warriors for a game played in the Bay Area?


----

## Data

The [`nbastatR`](http://asbcllc.com/nbastatR/index.html) package for `R` is a conventient API wrapper to pull rich player and game statistics from the NBA Stats API and Basketball-Reference (as well as other sources for non-NBA basketball leagues). You can install the package through github with devtools.

We'll load in the team and player reference tables, as well as game data for the past 3 seasons. We'll then transform the data structure somewhat in preparation for analysis. To starts, let's create dummy variables of the winners and losers for each game.


```{r dataload, echo=T, warning=FALSE, message=F}
library(nbastatR)
library(tidyverse)
library(lubridate)

today <- today(tzone = "EST")
historyCut <- year(today)-2

playerRef <- nba_players()
teamRef <- nba_teams()

teamCodes <- teamRef %>%
	dplyr::select(slugTeam, idTeam)

playerLogs <- game_logs(
  seasons = 2019,
  league = "NBA",
  result_types = "player",
  season_types = "Regular Season",
  nest_data = F,
  assign_to_environment = TRUE,
  return_message = TRUE
)

teamLogs <- game_logs(
  seasons = historyCut:year(today),
  league = "NBA",
  result_types = "team",
  season_types = "Regular Season",
  nest_data = F,
  assign_to_environment = TRUE,
  return_message = TRUE
) %>%
	dplyr::filter(locationGame=="H")


winners <- teamLogs %>%
    mutate(n=1) %>%
    pivot_wider(
      id_cols = idGame,
      names_from = slugTeamWinner,
      names_prefix = "win_",
      values_from = n,
      values_fill = list(n = 0)
    )


losers <- teamLogs %>%
    mutate(n=1) %>%
    pivot_wider(
      id_cols = idGame,
      names_from = slugTeamLoser,
      names_prefix = "lose_",
      values_from = n,
      values_fill = list(n = 0)
    )

winloss <- winners %>%
	left_join(losers)

```

----


## The Model

```{r modelling}

```


